%!TEX root = ../main.tex
%---------------------------------------------------------------------------------------------------
\subsection{Computation}\label{Appendix computation}\FloatBarrier
%---------------------------------------------------------------------------------------------------
Using the same computational implementation as \citetAppndx{Keane.1997}, we outline the immediate utility functions for each of the five alternatives. We first focus on their common structure and then present their parameterization. We also provide the economic motivation for their specification.\\

\noindent We follow individuals over their working life from age 16 until retirement at age 65. Each decision period $t = 16, \dots, 65$  represents a school year. Individuals can select one of five alternatives from the set of admissible actions $a\in\mathcal{A}$. They can decide to either work in a blue-collar or white-collar occupation ($a = 1, 2$), serve in the military $(a = 3)$, attend school $(a = 4)$, or stay at home $(a = 5)$.\\

\noindent Individuals differ with respect to their initial level of completed schooling $h_{16}$, and they possess one of four $\mathcal{J} = \{1, \hdots, 4\}$ alternative-specific skill endowments $\bm{e} = \left(e_{j,a}\right)_{\mathcal{J} \times \mathcal{A}}$.\\

\noindent The immediate utility $u_a(\cdot)$ of each alternative consists of a non-pecuniary utility $\zeta_a(\cdot)$ and, at least for the working alternatives, an additional wage component $w_a(\cdot)$, both of which depend on the level of human capital as measured by their occupation-specific work experience $\bm{k}_t = \left(k_{a,t}\right)_{a\in\{1, 2, 3\}}$, years of completed schooling $h_t$, and alternative-specific skill endowment $\bm{e}$. The immediate utility functions are influenced by last-period choices $a_{t -1}$ and alternative-specific productivity shocks $\bm{\epsilon}_t = \left(\epsilon_{a,t}\right)_{a\in\mathcal{A}}$ as well. Their general form is given by:
%
\begin{align*}
u_a(\cdot) =
\begin{cases}
    \zeta_a(\bm{k}_t, h_t, t, a_{t -1})  + w_a(\bm{k}_t, h_t, t, a_{t -1}, e_{j, a}, \epsilon_{a,t})                & \text{if}\, a \in \{1, 2, 3\}  \\
    \zeta_a(\bm{k}_t, h_t, t, a_{t-1}, e_{j,a}, \epsilon_{a,t})                                                  &  \text{if}\, a \in \{4, 5\}.
\end{cases}
\end{align*}
%
Work experience $\bm{k}_t$  and years of completed schooling $h_t$ evolve deterministically:
%
\begin{align*}
	k_{a,t+1} & = k_{a,t} + \ind[a_t = a]  \qquad \text{if}\, a \in \{1, 2, 3\} \\
	h_{t + 1\phantom{,a}} & = h_{t\phantom{,a}} +   \ind[a_t = 4].
\end{align*}
%
\noindent The productivity shocks are uncorrelated across time and follow a multivariate normal distribution with mean $\bm{0}$ and covariance matrix $\bm{\Sigma}$. Given the structure of the utility functions and the distribution of the shocks, the state at time $t$ is $s_t = \{\bm{k}_t, h_t, t, a_{t -1}, \bm{e},\bm{\epsilon}_t\}$.\\

\noindent Empirical and theoretical research from specialized disciplines within economics informs the exact specification of $u_a(\cdot)$. We now discuss each of its components in detail.
%-------------------------------------------------------------------------------
\subsubsection*{Non-pecuniary utility}
%-------------------------------------------------------------------------------
We begin by presenting the parameterization of the non-pecuniary utility for all five alternatives.
%-------------------------------------------------------------------------------
\paragraph{Blue-collar}
%-------------------------------------------------------------------------------
Equation (\ref{eq:NonWageBLueCollar}) shows the parameterization of the non-pecuniary utility from working in a blue-collar occupation:
%
\begin{align}\label{eq:NonWageBLueCollar}
\zeta_{1}(\bm{k}_t, h_t, a_{t-1})  = \alpha_1  &+ c_{1,1} \cdot \ind[a_{t-1} \neq 1] + c_{1,2} \cdot \ind[k_{1,t} = 0] \\ \nonumber
                            & + \vartheta_1 \cdot \ind[h_t \geq 12] + \vartheta_2 \cdot \ind[h_t \geq 16] + \vartheta_3 \cdot \ind[k_{3,t} = 1].
\end{align}
%
A constant $\alpha_1$ captures the net monetary equivalent of on-the-job amenities. Non-pecuniary utility includes mobility and search costs $c_{1,1}$, which are higher for individuals who had previously never worked in a blue-collar occupation, $c_{1,2}$, and captures returns from high school, $\vartheta_1$, and college degrees, $\vartheta_2$. Additionally, there is a detrimental effect of prematurely leaving the military after one year, $\vartheta_3$.
%-------------------------------------------------------------------------------
\paragraph{White-collar}
%-------------------------------------------------------------------------------
The non-pecuniary utility from working in a white-collar occupation is specified analogously. Equation (\ref{eq:UtilityWhiteCollar}) shows its parameterization:
%
\begin{align}\label{eq:UtilityWhiteCollar}
\zeta_{2}( \bm{k}_t, h_t, a_{t-1} ) = \,\alpha_2 & + c_{2,1} \cdot \ind[a_{t-1} \neq 2] + c_{2,2} \cdot \ind[k_{2,t} = 0]\\\nonumber
                            & + \vartheta_1 \cdot \ind[h_t \geq 12] + \vartheta_2 \cdot \ind[h_t \geq 16] + \vartheta_3 \cdot \ind[k_{3,t} = 1].
\end{align}
%-------------------------------------------------------------------------------
\paragraph{Military}
%-------------------------------------------------------------------------------
Equation (\ref{eq:UtilityMilitary}) shows the parameterization of the non-pecuniary utility from working in the military:
%
\begin{align}\label{eq:UtilityMilitary}
\zeta_{3}( k_{3.t}, h_t)  = \,& c_{3,2} \cdot \ind[k_{3,t} = 0]+ \vartheta_1 \cdot \ind[h_t \geq 12] + \vartheta_2 \cdot \ind[h_t \geq 16].
\end{align}
%
Although search costs $c_{3, 1} = 0$ are absent, there is a mobility cost if an individual has never previously served in the military, $c_{3,2}$. Individuals still experience a non-pecuniary utility from completing high school, $\vartheta_1$, and college, $\vartheta_2$.
%-------------------------------------------------------------------------------
\paragraph{School}
%-------------------------------------------------------------------------------
Equation (\ref{eq:UtilitySchooling}) shows the parameterization of the non-pecuniary utility from schooling:
%
\begin{align}\label{eq:UtilitySchooling}
	\zeta_4(k_{3,t}, h_t, t, a_{t-1}, e_{j,4}, \epsilon_{4,t})  = e_{j,4} & + \beta_{tc_1} \cdot \ind[h_t \geq 12] + \beta_{tc_2} \cdot \ind[h_t \geq 16]   \\\nonumber
    							  & + \beta_{rc_1} \cdot \ind[a_{t-1} \neq 4, h_t < 12]  \\\nonumber
    							  & + \beta_{rc_2} \cdot \ind[a_{t-1} \neq 4, h_t \geq 12]+ \gamma_{4,4} \cdot t  \\\nonumber
     							  & + \gamma_{4,5} \cdot \ind[t < 18] + \vartheta_1 \cdot \ind[h_t \geq 12]  \\\nonumber
      							& + \vartheta_2 \cdot \ind[h_t \geq 16]+ \vartheta_3 \cdot \ind[k_{3,t} = 1] + \epsilon_{4,t}.
\end{align}
%
There are direct costs for pursuing higher education, which primarily take the form of college, $\beta_{tc_1}$, and graduate school tuition fees, $\beta_{tc_2}$. The decision to leave school is reversible, but entails adjustment costs that differ by schooling category ($\beta_{rc_1}, \beta_{rc_2}$). Education is defined by time spent in school, not by formal credentials acquired. Once individuals reach a certain amount of schooling, they acquire a degree. There is no uncertainty about grade completion \citepAppndx{Altonji.1993} and no part-time enrollment. Individuals value the completion of high school and college ($\vartheta_1, \vartheta_2$).
%-------------------------------------------------------------------------------
\paragraph{Home}
%-------------------------------------------------------------------------------
Equation (\ref{eq:UtilityHome}) shows the parameterization of the non-pecuniary utility from staying at home:
%
\begin{align}\label{eq:UtilityHome}
	\zeta_5(k_{3,t}, h_t, t, e_{j,5}, \epsilon_{5,1}) =  e_{j,5} & + \gamma_{5,4} \cdot \ind[18 \leq t \leq 20] + \gamma_{5,5} \cdot \ind[t \geq 21] \\ \nonumber
    							   & +\vartheta_{1} \cdot \ind[h_t \geq 12] + \vartheta_{2} \cdot \ind[h_t \geq 16] \\ \nonumber
    							   & +  \vartheta_3 \cdot \ind[k_{3,t} = 1]+ \epsilon_{5,t}.
\end{align}
%
Staying at home as a young adult, $\gamma_{5, 4}$, is less stigmatized than doing so as an older individual, $\gamma_{5,5}$. Possessing a degree $(\vartheta_1, \vartheta_2)$ or leaving the military prematurely, $\vartheta_3$, influences the immediate utility as well.
%-------------------------------------------------------------------------------
\subsubsection*{Wage component}
%-------------------------------------------------------------------------------
The wage component $w_{a}(\cdot)$ for the working alternatives is given by the product of the market-equilibrium rental price $r_{a}$ and an occupation-specific skill level $x_{a}(\cdot)$. The latter is determined by the overall level of human capital:
%
\begin{align*}
w_{a}(\cdot) = r_{a} \, x_{a}(\cdot).
\end{align*}
%
This specification leads to a standard logarithmic wage equation in which the constant term is the skill rental price $\ln(r_{a})$ and wages follow a log-normal distribution.\\

\noindent The occupation-specific skill level $x_{a}(\cdot)$ is determined by a skill production function, which includes a deterministic component $\Gamma_a(\cdot)$ and a multiplicative stochastic productivity shock $\epsilon_{a,t}$:
%
\begin{align}
    x_{a}(\bm{k}_t, h_t, t, a_{t-1}, e_{j, a}, \epsilon_{a,t}) & = \exp \big( \Gamma_{a}(\bm{k}_t,  h_t, t, a_{t-1}, e_{j,a}) \cdot \epsilon_{a,t} \big). \nonumber
\end{align}
%-------------------------------------------------------------------------------
\paragraph{Blue-collar}
%-------------------------------------------------------------------------------
Equation (\ref{eq:SkillLevelBlueCollar}) shows the parameterization of the deterministic component of the skill production function:
%
\begin{align}\label{eq:SkillLevelBlueCollar}
    \Gamma_1(\bm{k}_t, h_t, t, a_{t-1}, e_{j, 1}) = e_{j,1} & + \beta_{1,1} \cdot h_t + \beta_{1, 2} \cdot \ind[h_t \geq 12] \\ \nonumber
                                  & + \beta_{1,3} \cdot \ind[h_t\geq 16] + \gamma_{1, 1} \cdot  k_{1,t} + \gamma_{1,2} \cdot  (k_{1,t})^2 \\ \nonumber
                                & + \gamma_{1,3} \cdot  \ind[k_{1,t} > 0] + \gamma_{1,4} \cdot  t + \gamma_{1,5} \cdot \ind[t < 18]\\ \nonumber
                                  & + \gamma_{1,6} \cdot \ind[a_{t-1} = 1] + \gamma_{1,7} \cdot  k_{2,t} + \gamma_{1,8} \cdot  k_{3,t}. \nonumber
\end{align}
%

\noindent There are several notable features. The first part of the skill production function is motivated by \citetAppndx{Mincer.1974} and, hence, linear in years of completed schooling, $\beta_{1,1}$, quadratic in experience ($\gamma_{1,1}, \gamma_{1,2}$), and separable between the two of them. There are so-called sheep-skin effects \citepAppndx{Hungerford.1987, Jaeger.1996} associated with completing a high school, $\beta_{1,2}$, or graduate education, $\beta_{1,3}$, which capture the impact of completing a degree beyond the associated years of schooling. There is also a first-year blue-collar experience effect $\gamma_{1,3}$. Additionally, job skills depreciate for blue-collar workers, who were unemployed in the previous period, $\gamma_{1,6}$. All forms of work experience ($\gamma_{1,7}, \gamma_{1,8}$) are transferable.
%-------------------------------------------------------------------------------
\paragraph{White-collar}
%-------------------------------------------------------------------------------
The wage component from working in a white-collar occupation is specified analogously. Equation (\ref{eq:SkillLevelWhiteCollar}) shows the parameterization of the deterministic component of the skill production function:
%
\begin{align}\label{eq:SkillLevelWhiteCollar}
    \Gamma_2(\bm{k}_t, h_t, t, a_{t-1}, e_{j,2}) = e_{j,2} & + \beta_{2,1} \cdot h_t + \beta_{2, 2} \cdot \ind[h_t \geq 12]  \\\nonumber
    							 & + \beta_{2,3} \cdot \ind[h_t\geq 16]+ \gamma_{2, 1} \cdot  k_{2,t} + \gamma_{2,2} \cdot  (k_{2,t})^2  \\\nonumber
                                   & + \gamma_{2,3} \cdot  \ind[k_{2,t} > 0]+ \gamma_{2,4} \cdot  t + \gamma_{2,5} \cdot \ind[t < 18] \\\nonumber
                                  & + \gamma_{2,6} \cdot  \ind[a_{t-1} = 2]  + \gamma_{2,7} \cdot  k_{1,t} + \gamma_{2,8} \cdot  k_{3,t}.
\end{align}
%-------------------------------------------------------------------------------
\paragraph{Military}
%-------------------------------------------------------------------------------
Equation (\ref{eq:SkillLevelMilitary}) shows the parameterization of the deterministic component of the skill production function:
%
\begin{align}\label{eq:SkillLevelMilitary}
    \Gamma_3( k_{3,t}, h_t, t, e_{j,3}) = e_{j,3} & + \beta_{3,1} \cdot h_t  \\\nonumber
	               \nonumber &+ \gamma_{3,1} \cdot  k_{3,t} + \gamma_{3,2} \cdot (k_{3,t})^2 + \gamma_{3,3} \cdot \ind[k_{3,t} > 0]\\\nonumber
									 & + \gamma_{3,4} \cdot t + \gamma_{3,5} \cdot \ind[t < 18].
\end{align}
%
Unlike the civilian sector, there are no sheep-skin effects from completing military training, ($\beta_{3,2} = \beta_{3,3}= 0$). Furthermore, the previous occupational choice has no influence ($\gamma_{3,6}= 0$), and any experience other than military is non-transferable ($\gamma_{3,7} = \gamma_{3,8} = 0$).

\begin{Remark} Our parameterization for the immediate utility of serving in the military differs from \citetAppndx{Keane.1997}, as we remain unsure about their exact specification. The authors state in Footnote 31 (p.\,498) that the constant for the non-pecuniary utility $\alpha_{3,t}$ depends on age. However, we are unable to determine the precise nature of the relationship. Equation (C3) (p.\,521) also indicates no productivity shock $\epsilon_{a,t}$ in the wage component. Table~7 (p.\,500) reports such estimates.
\end{Remark}

\noindent Table \ref{Model parameters} presents an overview of the model's parameters.

\input{../material/tab-model-parameters.tex}
