% !TEX root = ../main.tex
\title{Structural models for policy-making\thanks{Corresponding author: Philipp Eisenhauer, peisenha@uni-bonn.de. Philipp Eisenhauer and Lena Janys are both funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany's Excellence Strategy - EXC 2126/1- 390838866 and the TRA Modelling (University of Bonn) as part of the Excellence Strategy of the federal and state governments. Jano\'s Gabler is grateful for financial support by the German Research Foundation (DFG) through CRC-TR 224 (Project C01) and funding by IZA Institute of Labor Economics. Lena Janys is funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany's Excellence Strategy - EXC 2047/1 - 390685813. Philipp Eisenhauer was funded by a postdoctoral fellowship by the AXA Research Fund. We thank Tim Mensinger for his help in the early stages of the project. We thank Max Blesch, Joachim Freyberger, Annica Gehlen, Daniel Harenberg, Ken Judd, Gregor Reich, J\"org Stoye, and Rafael Suchy for numerous helpful discussions. We thank Michael Keane and Kenneth Wolpin for providing the dataset used in our analysis. We thank Annica Gehlen and Emily Schwab for their outstanding research assistance. We are grateful to the Social Sciences Computing Service (SSCS) at the University of Chicago for the permission to use their computational resources. Eisenhauer: University of Bonn, peisenha@uni-bonn.de; Gabler: University of Bonn and IZA, janos.gabler@uni-bonn.de, Janys: University of Bonn, ljanys@uni-bonn.de}}
\subtitle{Coping with parametric uncertainty}
\author{Philipp Eisenhauer \& Jano\'s Gabler \& Lena Janys \& Christopher Walsh\\ University of Bonn\vspace{0.5cm}}
\date{\today}
